---
title: "packagemetrics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
```

## rOpenSci Unconference 2017 project repo

This repo is for a package to obtain a collection of metrics on R packages which are intended to inform the decision which package you choose for a certain task.

This is based on issues [#78 Avoiding redundant / overlapping packages](https://github.com/ropensci/unconf17/issues/78) and [#69 A framework for reproducible tables](https://github.com/ropensci/unconf17/issues/69) for the unconf 17.

## Team members:

Lori Shepherd, Hannah Frick, William Ampeh, Erin Grand, Sam Firke, Becca Krouse

## Goals:

Develop a process for assessing and comparing a set of packages. Action items include:

- Develop a set of objective metrics based off of CRAN/GitHub/StackOverflow
- Develop a process for subjective assessment
- Apply to a set of table-making packages

## Collection of ideas for metrics:

### CRAN

- Does it have a public repo?
- Does it have an issue tracker or similar?
- Dependencies
    - Reverse dependencies - are others depending on this?
    - Dependencies - is it built on a solid stack?
- Is the package being maintained?
    - Date of first release?
    - Date of latest release
- Does it have a vignette (how many?)


### GitHub

- Is the package being maintained?
    - Last commit?
    - How many users have committed in the last [interval of time]?
    - Issue tracker?
    - When was the last issue closed?
- Tests    
    - Does the package have tests?
    - What's the test coverage?
    - Does it have continuous integration?
- Analysis of help pages 
    - Keywords
    - Similarity of multiple pkgs
    - What % of function man pages have an example?
- What packages are commonly co-mentioned with it?  Could be in its own documentation/vignette or on GitHub scripts where it is mentioned


### StackOverflow

- Popularity in stackoverflow answers 
- What packages are commonly co-mentioned with it?

### Other

- User-friendliness - is it modular, “tidy”, clean syntax, coherent theory, not flexible, not trying to do too much
- Are there any/many blog posts on/which mention a pkg?

## Installation
```
devtools::install_github("ropenscilabs/packagemetrics")
```

## Use

### Here's an example using a case study of popular table-making packages.  

```{r, echo=TRUE}
library(packagemetrics)

pkg_df <- package_list_metrics(table_packages)
```

```{r, echo=FALSE}
library(dplyr)
library(formattable)
pd <- pkg_df %>% 
  select(package, published, dl_last_month, stars,
         tidyverse_happy, has_tests, has_vignette_build, 
         last_commit, last_issue_closed, contributors, 
         depends_count, reverse_count) %>% 
  mutate(last_commit = round(last_commit, 1),
         last_issue_closed = round(last_issue_closed, 1)) %>% 
  rename(vignette = has_vignette_build) 
pd[is.na(pd)] <- ""

t <- formattable(pd, list(
    package=formatter("span",
                      style = x ~ style(font.weight = "bold")),
    contributors = color_tile("white","#1CC2E3"),
    depends_count = color_tile("white", "#1CC2E3"),
    reverse_count = color_tile("white", "#1CC2E3"),
    tidyverse_happy = formatter("span",
                                style = x ~ style(color = ifelse(x, "purple","white")),
                                x ~ icontext(ifelse(x, "glass","glass"))),
    vignette = formatter("span",
                         style = x ~ style(color = ifelse(x, "green","white")),
                         x ~ icontext(ifelse(x, "ok","ok"))),
    has_tests =  formatter("span",
                           style = x ~ style(color = ifelse(x, "green","red")),
                           x ~ icontext(ifelse(x, "ok","remove"))),
    dl_last_month = color_bar("#56A33E"),
    forks = color_tile("white", "#56A33E"),
    stars = color_tile("white", "#56A33E"),
    last_commit = color_tile("#F06B13","white", na.rm=T),
    last_issue_closed = color_tile("#F06B13","white", na.rm=T)
  ))

w <- formattable::as.htmlwidget(t,width=1200, height=NULL)
path <- htmltools::html_print(w, background = "white", viewer = NULL)
url <- paste0("file:///", gsub("\\\\", "/", normalizePath(path)))
webshot::webshot(url,
                 file = "inst/image/table.png",
                 selector = ".formattable_widget",
                 delay = 0.2)

```

## Requirements

```packagemetrics``` requires R version >= 3.4.0.
