---
title: "tableSample"
author: "rTables-Group"
date: "May 26, 2017"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
***

## Selected Tables packages with sample code

As a proof of concept, below is the List of 5 Table packages we considered today:


### 1: Arsenal

An Arsenal of 'R' functions for large-scale statistical summaries, which are streamlined to work within the latest reporting tools in 'R' and 'RStudio' and which use formulas and versatile summary statistics for summary tables and model.

[Documentation](https://www.rdocumentation.org/packages/arsenal/versions/0.2.0)

#### Example

```{r arsenal, warning=FALSE, message=FALSE}
#install.packages("arsenal")
library(arsenal)
library(dplyr)

data(mockstudy)
tab1 <- tableby(arm ~ sex + age, data=mockstudy)
mylabels <- list(sex = "SEX", age ="Age, yrs")
summary(tab1, labelTranslations = mylabels, text=TRUE)


#freqlist() is a function to approximate the output from SAS's PROC FREQ procedure when using the /list option of the TABLE statement.
# load mockstudy data
data(mockstudy)
tab.ex <- table(mockstudy[, c("arm", "sex", "mdquality.s")], useNA = "ifany")
noby <- freqlist(tab.ex, na.options = "include")
summary(noby)


```


***
***

### 2: condformat

Prints a data frame with cells formatted according to several rules or criteria. It is integrated with the RStudio Viewer or a web browser, and it supports knitr and rmarkdown outputs using both HTML and PDF output formats.

[Documentation](https://cran.r-project.org/web/packages/condformat/vignettes/introduction.html)

#### Example

```{r condformat, warning=FALSE, message=FALSE}
data(iris)
#install.packages("condformat")
library(condformat)
condformat(iris[c(1:5,70:75, 120:125),]) +
  rule_fill_discrete(Species) + 
  rule_fill_discrete(Sepal.Width, Sepal.Length,
                     expression = Sepal.Width > Sepal.Length - 2.25,
                     colours = c("TRUE" = "#7D00FF")) + 
  rule_fill_gradient2(Petal.Length)

```


***
***


### 3: desctable

A comprehensive descriptive and comparative tables generator for R.

[Documentation](https://cran.r-project.org/web/packages/desctable/vignettes/desctable.html)

#### Example

```{r desctable, warning=FALSE, message=FALSE}
#install.packages("desctable")
library(desctable)

iris %>%
  desctable

#The object produced by desctable is in fact a list of data.frames, with a "desctable" class. Methods for reduction to a simple dataframe (as.data.frame, automatically used for printing), conversion to markdown (pander

mtcars %>%
  desctable %>%
  datatable
```

***
***

### 4: expss

Provides tabulation functions with support of SPSS-style labels, multiple / nested banners, weights and multiple-response variables.

[Documentation](https://cran.r-project.org/web/packages/expss/vignettes/TablesWithLabelsInR.html)

#### Example

```{r expss, warning=FALSE, message=FALSE}
#install.packages("expss")
library(expss)
data(mtcars)
mtcars = apply_labels(mtcars,
                      mpg = "Miles/(US) gallon",
                      cyl = "Number of cylinders",
                      disp = "Displacement (cu.in.)",
                      hp = "Gross horsepower",
                      drat = "Rear axle ratio",
                      wt = "Weight (1000 lbs)",
                      qsec = "1/4 mile time",
                      vs = "Engine",
                      vs = c("V-engine" = 0,
                             "Straight engine" = 1),
                      am = "Transmission",
                      am = c("Automatic" = 0,
                             "Manual"=1),
                      gear = "Number of forward gears",
                      carb = "Number of carburetors"
)

#For quick cross-tabulation there are fre and cro family of function. For simplicity we demonstrate here only cro_cpct which caluclates column percent. Documentation for other functions, such as cro_cases for counts, cro_rpct for row percent, cro_tpct for table percent and cro_fun for custom summary functions can be seen by typing ?cro and ?cro_fun in the console.

# 'cro' examples
# multiple banners
mtcars %>% 
    calculate(cro_cpct(cyl, list(total(), am, vs))) %>% 
    htmlTable(caption = "Table with multiple banners (column %).")
```

***
***

### 5: formattable

Designed for applying formatting on vectors and data frames to make data presentation easier, richer, more flexible and hopefully convey more information. 

[Documentation](https://cran.r-project.org/web/packages/formattable/vignettes/introduction.html)

#### Example

```{r formattable, warning=FALSE, message=FALSE}
#install.packages("formattable")
library(formattable)
p <- percent(c(0.1, 0.02, 0.03, 0.12))      # display values in percent format
p

#works with arithmetic operations and other common functions and preserves its formatting.
p + 0.05
```

***
***


### 6: kableExtra

Designed to help build common complex tables and manipulate table style.  Comes with hoverable html tables, multi-column header and footnotes capabilities.

[Documentation](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

#### Example

```{r kableExtra, warning=FALSE, message=FALSE}
#install.packages(knitr)
#install.packages("kableExtra")
library(knitr)
library(kableExtra)

dt <- mtcars[1:5, 1:6]

#Basic HTML Table
kable(dt)

#HTML hoverable tables 
kable(dt, "html") %>%
   kable_styling(bootstrap_options = c("striped", "hover"))

# Add Extra Header Rows
dt %>%
  kable("html") %>%
  kable_styling("striped") %>%
  add_header_above(c(" " = 1, "Group 1" = 2, "Group 2" = 2, "Group 3" = 2))

# Add footnote
dt %>%
  kable("html", caption = "Demo Table[note]") %>%
  kable_styling("striped") %>%
  add_header_above(c(" ", "Group 1[note]" = 3, "Group 2[note]" = 3)) %>%
  add_footnote(c("This table is from mtcars", 
                 "Group 1 contains mpg, cyl and disp", 
                 "Group 2 contains hp, drat and wt"), 
               notation = "symbol")
```

***
***

### 7: tableby 

Provides similar functionality as the SAS macros %table and %summary. 

[Documentation](https://cran.r-project.org/web/packages/arsenal/vignettes/tableby.html)

#### Example

```{r tableby, warning=FALSE, message=FALSE}
# install.packages("arsenal")
# install.packages("knitr")
# install.packages("survival")

require(arsenal)
require(knitr)
require(survival)
data(mockstudy) ##load data
#dim(mockstudy)  ##look at how many subjects and variables are in the dataset 

#Create a table stratified by treament arm, use a formula statement to specify the variables that you want summarized. The example below uses age (a continuous variable) and sex (a factor).

tab1 <- tableby(arm ~ sex + age, data=mockstudy)

summary(tab1, text=TRUE)

```

***
***

### 8: Huxtable 

A package for creating HTML and LaTeX tables. It provides similar functionality to xtable, but more, with a simpler interface.

[Documentation](https://hughjonesd.github.io/huxtable/)

#### Example

```{r huxtable, warning=FALSE, message=FALSE}
#install.packages("psych")
#install.packages("dplyr")
#install.packages("magrittr")
#install.packages("huxtable")

library(psych)
library(dplyr)
library(magrittr)
library(huxtable)

data(attitude)
att_corr <- corr.test(as.matrix(attitude))

as_hux(att_corr$r)                                                             %>%
      set_background_color(where(att_corr$p < 0.05), 'yellow')                 %>%
      set_background_color(where(att_corr$p < 0.01), 'orange')                 %>%
      set_background_color(where(row(att_corr$p) == col(att_corr$p)), 'white') %>%
      huxtable::add_rownames(colname = '')                                     %>%
      huxtable::add_colnames()                                                 %>%
      set_all_borders(1)

```

***
***

### 9: Others

List of tabling packages to be considered later


(1) [ascii](https://cran.r-project.org/web/packages/ascii/index.html)
(2) [ezsummary](https://cran.r-project.org/web/packages/ezsummary/index.html) (https://cran.r-project.org/web/packages/ezsummary/README.html)
(3) [gmodels::CrossTable()](https://www.rdocumentation.org/packages/gmodels/versions/2.16.2/topics/CrossTable)
(3) [Gtable with grid.draw, grid.table and tableGrob](https://cran.r-project.org/web/packages/gridExtra/vignettes/tableGrob.html)
(4) [Hmisc::latex](https://www.rdocumentation.org/packages/Hmisc/versions/4.0-3)
(5) [htmlTable](https://cran.r-project.org/web/packages/htmlTable/vignettes/tables.html)
(6) [janitor](https://cran.r-project.org/web/packages/janitor/vignettes/introduction.html)
(7) [kable](https://www.rdocumentation.org/packages/knitr/versions/1.16/topics/kable)
(8) [pander](https://www.r-project.org/nosvn/pandoc/pander.html)
(9) [pixiedust](https://cran.r-project.org/web/packages/pixiedust/vignettes/advancedMagic.html)
(10) [ReporteRs](http://www.sthda.com/english/wiki/create-and-format-word-documents-using-r-software-and-reporters-package)
(11) [stargazer](https://www.r-statistics.com/2013/01/stargazer-package-for-beautiful-latex-tables-from-r-statistical-models-output/)
(12) [Tableone](https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html)
(13) [tables](http://www.cyclismo.org/tutorial/R/tables.html)
(14) [Tabular](https://cran.r-project.org/web/packages/tables/vignettes/tables.pdf)
(15) [tangram](https://github.com/spgarbet/tangram/blob/master/vignettes/example.Rmd) 
(16) [texreg](https://diffuseprior.wordpress.com/2013/01/20/texreg-a-package-for-beautiful-and-easily-customizable-latex-regression-tables-from-r/)
(17) [xtable](https://cran.r-project.org/web/packages/xtable/vignettes/OtherPackagesGallery.pdf)
(18) [Ztable](https://cran.r-project.org/web/packages/ztable/vignettes/ztable.html)




